A Review of Neutron Imaging

### 2020-10-10 neutron-imaging/neutron.tex
WIP 10Oct2020

Neutrons are best shielded by material rich in protons like Iron, water, 
paraffin, concrete and graphite, in decreasing order of attenuation length
\cite{turner2008atoms}.
\section{Propagation of neutron beams}\label{s7}
\section{Neutron diffraction by crystals}\label{s8}
\section{2-d neutron imaging}\label{s9}
\section{Neutron tomography}\label{s10}
\section{Imaging of biologically important structures}\label{s11}
\section{Larmor precession based imaging}\label{s12}
\section{Polarization based imaging}\label{s13}
\section{Conclusion}\label{s14}
\bibliography{neutron}
\bibliographystyle{plain}
\end{document}

>>>>

Neutrons are best shielded by material rich in protons like Iron, water, 
paraffin, concrete and graphite, in decreasing order of attenuation length
\cite{turner2008atoms}.

\section{Propagation of neutron beams}\label{s7}
Typical neutron sources have size of the order of $10$ cm in all directions.
They are enclosed in moderators to produce a quasi-Maxwell distribution
\cite{anderson2008neutron}. Their brightness is defined as the flux of neutrons
per solid angle per unit wavelength. The flux itself is the number of neutrons
crossing a unit area. One can treat the neutrons using the principles of
geometrical optics. The phase space of analytical mechanics can be carried over
to optics and is called etendue. Consider an infinitesimal surface element 
$dS$ with normal $\hat{n}$ pointing in a medium of refractive index $n$. If
the surface $dS$ is crossed by neutrons confined to a solid angle $d\Omega$
then the etendue of neutron beam is
\begin{equation}\label{s7e1}
dG = n^2\hat{n}\cdot\hat{n}_b dSd\Omega,
\end{equation}
where $\hat{n_b}$ is the normal along the axis of the cone of $d\Omega$ 
\cite{markvart2007thermodynamics}. The way the volume of the phase space is
conserved, one can also conclude that the volume of an etendue is conserved. 
The optical analogue of Liouville's theorem places a limit on the efficiency
of a neutron source \cite{anderson2008neutron}, 
\cite{markvart2007thermodynamics}. If one want to focus a neutron beam from
a source of size $w_1$ on a spot of size $w_2 < w_1$ then one must allow the
beam to diverge in directions normal to the line joining the source and the
spot.

Neutron beams can also be transported away from the source using wave guides.
They are channels coated with Nickel which has a critical angle of $0.10^\circ$
per $\AA$ neutron wavelength. Where higher critical angles than Nickel are 
needed super-mirrors are used. Critical angles for super-mirrors can be $4$
to $6$ times more than Nickel. Oftentimes it is beneficial to carry neutrons
away from the source to prevent their contamination with fast neutrons which
are also emitted from the source. Allowing the neutrons a longer flight time
through wave guides also improves the resolution of pulsed-source instruments.
The latter quantity is proportional to $\Delta t/t$, where $t$ is the total
time of flight and $\Delta t$ is the time width of the pulse in the moderator.

A neutron source produces neutrons with a wide range of wavelengths. A device
that allows us to select one among them is called a monochromator. Crystal
monochromators are based on Bragg's law
\begin{equation}\label{s7e2}
n\lambda = 2d\sin\theta,
\end{equation}
where $n$ is the `order' of diffraction, $\lambda$ is the wavelength of 
neutrons, $d$ is the distance between the lattice planes and $\theta$ is the
angle between the crystal and the incident beam. (The angle of incidence is
$\pi/2 - \theta$.) Once a crystal is chosen, $d$ is fixed. One can select
the wavelength of choice by changing $\theta$. The intensity of the diffracted
beam reduces for higher order of diffractions. Therefore, monochromators us
the first order beam.

The direction of propagation of neutron beams can be changed by refraction.
The refractive index of neutron beams is given by the Fermi thin-slab formula
\cite{warner1985neutron}
\begin{equation}\label{s7e3}
n = 1 - \frac{2\pi\rho b}{k_0^2},
\end{equation}
where $b$ is the neutron scattering length, $\rho$ is the number density of
the scattering nuclei and $k_0$ is the propagation constant in free space.
A refractive index lesser than $1$ indicates that one must use a concave lens
to focus a neutron beam. Although a neutron is electrically neutral, it does
have spin. Therefore, it can be deflected using a magnetic field. If 
$\vec{r}(t)$ is the trajectory of a neutron beam then its deflection in a
magnetic field $\vec{B}$ is given by
\begin{equation}\label{s7e4}
\frac{d^2\vec{r}}{dt^2} = \pm\frac{\mu}{m}\nabla|\vec{B}|,
\end{equation}
where $\mu$ is the neutron's magnetic moment and $m$ is its mass.

\section{Neutron diffraction by crystals}\label{s8}
\section{2-d neutron imaging}\label{s9}
\section{Neutron tomography}\label{s10}
\section{Imaging of biologically important structures}\label{s11}
\section{Larmor precession based imaging}\label{s12}
\section{Polarization based imaging}\label{s13}
\section{Conclusion}\label{s14}
\bibliography{neutron}
\bibliographystyle{plain}
\end{document}
